A tutorial of Solving the defined integral problem with OpenMp and Open MPI

### 2023-07-11 documentation/article03/article.tex
Open MPI implementation

\begin{document}
	\title{A tutorial of Solving the defined integral problem with OpenMp}
	\author{André Furlan - UNESP - Universidade Estadual Paulista "Júlio de Mesquita Filho"}
	\date{2023-06-01}
	\maketitle

	\begin{abstract}
		The Trapezoid rule algorithm used in this tutorial is \textbf{not} the most fast method to calculate an interval defined integral but is a good example of how to make programs runs quicker by using OpenMp for parallelism. Hopefully it will help the understanding of how this library works and how to use it. All sources can be accessed at \href{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}
	\end{abstract}

	\begin{IEEEkeywords}
		OpenMp, Paralelism, C++, C, Tutorial
	\end{IEEEkeywords}

	\section{Introduction}
	\par In High Performance Computation (HPC) the goal (other than solve a problem) is to use most of the computation power of the machines involved. Most of the time this means use all processors an its cores at maximum, avoiding idle intervals.
	\par By using the OpenMp \cite{openmp08} libraries it is possible, using \textbf{\#pragma} compiler directives, turn a sequential program into a parallel one. The problem in focus here is the defined interval integral calculation. To solve it the program must calculate the area bellow the plot by summing up the area of several trapezoids \cite{hildebrand1987introduction}. The more trapezoids fitted, the more precise are the values obtained. And it is here that the parallelism enters: By dividing the work of area calculation in lots of threads its possible to diminish the effective time of execution.
	\section{The trapezoid rule}
	\par Be $a$ a point that marks the start of a interval in which the calculation must be done and $b$ its end, than an \textbf{approximation} of the integral value is given by equation \ref{eq:firstApproximation}. In figure \ref{fig:trapezoidalruleillustration} the basic idea is illustrated.\newline

@@

		\par Then, for the impatient ones, here is the complete code that will be discussed:
		\lstinputlisting[language=C++]{code/integral.cpp}

>>>>

\begin{document}
	\title{A tutorial of Solving the defined integral problem with OpenMp and Open MPI}
	\author{André Furlan - UNESP - Universidade Estadual Paulista "Júlio de Mesquita Filho"}
	\date{2023-06-01}
	\maketitle

	\begin{abstract}
		The Trapezoid rule algorithm used in this tutorial is \textbf{not} the most fast method to calculate an interval defined integral but is a good example of how to make programs runs quicker by using OpenMp or Open MPI for parallelism. Hopefully it will help the understanding of how these libraries works and how to use them. All sources can be accessed at \href{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}
	\end{abstract}

	\begin{IEEEkeywords}
		OpenMp, Open MPI, Paralelism, C++, C, Tutorial
	\end{IEEEkeywords}

	\section{Introduction}
	\par In High Performance Computation (HPC) the goal (other than solve a problem) is to use most of the computation power of the machines involved. Most of the time this means use all processors an its cores at maximum, avoiding idle intervals.
	\par By using the OpenMp \cite{openmp08} library it is possible by using \textbf{\#pragma} compiler directives, turn a sequential program into a parallel one.

	\par When using the Open MPI library, \cite{mpi40} it is necessary to use the "MPI\_" prefixed commands. The "MPI\_" prefix is a convention used in Open MPI implementations, but it is not required by the MPI standard.

	\par The problem in focus here is the defined interval integral calculation. To solve it the program must calculate the area bellow the plot by summing up the area of several trapezoids \cite{hildebrand1987introduction}. The more trapezoids fitted, the more precise are the values obtained. And it is here that the parallelism enters: By dividing the work of area calculation in lots of threads its possible to diminish the effective time of execution.

	\section{The trapezoid rule}
	\par Be $a$ a point that marks the start of a interval in which the calculation must be done and $b$ its end, than an \textbf{approximation} of the integral value is given by equation \ref{eq:firstApproximation}. In figure \ref{fig:trapezoidalruleillustration} the basic idea is illustrated.\newline

@@

		\subsection{OpenMp implementation}	
			\lstinputlisting[language=C++]{code/integral.cpp}

		\subsection{Open MPI implementation}	
			\lstinputlisting[language=C++]{code/integral2.cpp}
